Constraints on the quasi-normal mode frequencies of the LIGO-Virgo signals by making full use of gravitational-wave modeling

### 2021-01-25 papers/intro_paper/intro_paper.tex
edits to intro<

\title{\ab{Constraints on the quasi-normal mode frequencies of the LIGO-Virgo signals by making full use of gravitational-wave modeling}}

@@

The LIGO Scientific Collaboration~\citep{lsc} and the Virgo
Collaboration~\citep{Virgo} have recently announced their catalogue of
gravitational-wave (GW) signals from the first
half of the third observing run (O3a)~\cite{Abbott:2020niy}. Combined
with the first and second observing-run catalogues~\cite{LIGOScientific:2018mvr}, the Advanced LIGO detectors at Hanford,
Washington and Livingston, Louisiana~\citep{aasi2015characterization},
and the Advanced Virgo detector in Cascina,
Italy~\citep{acernese2014advanced} have now detected \sout{more than} $50$ GW
events from the merger of compact objects like neutron stars and/or
black holes (BHs).\sout{, collectively called compact binary coalescences
(CBCs).} Alongside independent claims of
detections~\citep{nitz20191,nitz20202,2019PhRvD.100b3007Z,2020PhRvD.101h3030V,Venumadhav_2020},
\sout{this} \ab{these results} \sout{has} \ab{have} firmly established the field of GW astronomy, five years \sout{on
from} \ab{after} the first \sout{ever direct} detection of a GW \ab{passing through the} \sout{on} Earth, \ab{notably GW150914} 
\sout{on September 14,2015}~\citep{abbott2016observation}.

\ab{The observation of GWs has} had significant astrophysical and cosmological
implications~\citep{LSC_2016astroph,gw170817_mma,gw170817_joint,gw170817_hubble}. It
has also allowed us to \sout{make statements in} \ab{probe} fundamental
physics \sout{.Specifically, LIGO-Virgo's observations have allowed us to} and 
test predictions of Einstein's theory of General Relativity 
(GR)~\citep{GR} in the previously unexplored \ab{highly-dynamical, and strong field 
regime} \sout{of highly relativistic, strong-field regimes of
gravity}~\citep{LSC_2016grtests,GW170817_TGR,gwtc1_tgr}. In GR\sout{, a CBC
involving two BHs (}a binary black hole (BBH) system\sout{)} is described \sout{in} \ab{by}
three \sout{distinct} phases: an early \textit{inspiral}, where the two
compact objects spiral in \ab{loosing energy because of the emission of GWs} \sout{and \textit{plunge} due to a back-reaction
of GW-emission}, a \textit{merger} marked by the
formation of a\sout{n} \ab{common} apparent horizon~\citep{NRpaper}, and a \sout{late-time} 
\textit{ringdown}, \sout{where} \ab{during which} the newly formed remnant object settles down
to a \sout{stable} Kerr \ab{BH} \sout{state through the emission of} \ab{emitting} \sout{an exponentially damped} 
quasi-normal-modes (QNMs) \sout{spectrum of gravitational
radiation}~\citep{vishu,earlyqnmpapers} \ab{(i.e., damped oscillations with specific, discrete 
frequencies and decay times)}.

The LIGO-Virgo collaborations \ab{have released companion papers \cite{}} detailing their \sout{latest} results 
of tests of GR  \sout{using events} \ab{for GW150914} \cite{} and \ab{for several GW events of} 
the \sout{first} two transient catalogues (TC): GWTC-1\cite{}  and GWTC-2 \cite{}. 
The results include tests of GW generation and source dynamics, where bounds are placed on
parameterized deviations in the post-Newtonian \ab{(PN)} coefficients describing
the early inspiral~\citep{earlydevelopmentpapers}, and
phenomenological coefficients describing the intermediate (plunge) and
merger regimes of coalescence~\citep{TIGERmethodspapers}; tests of GW
propagation, which assume a generalized dispersion relation and place
upper bounds on the Compton wavelength and, consequently, the mass of
the graviton~\citep{gw170104,samajdar2017projected}, and tests of the
polarization of gravitational radiation using a
multi--GW-detector network~\citep{gw170814,isi2017probing}. The \ab{GWTC-1/2} paper\ab{s} also 
check\sout{s} for consistency between different portions of the signal using estimates
\sout{of} \ab{for the} predicted mass and spin of the remnant
object~\citep{Ghosh:2016xx,Ghosh:2017gfp,LSC_2016grtests}, and
consistency of the residuals with interferometric
noise~\citep{Ghonge:2020suv,gwtc1_tgr}. None of these tests report any
departure from the predictions of GR.

\ab{The first paper on tests of GR by the LIGO Collaboration~\cite{} provided us with the 
first measurement of the dominant damped-oscillation signal in the ringdown stage of a BBH 
coalescence.} The LIGO-Virgo O3a testing GR paper~\cite{} \ab{has recently reported} 
\sout{also details results on tests} \ab{a comprehensive analysis of the properties of the 
remnant object, including the ringdown stage, for tens of GW events.}   
\sout{BH ringdown and nature of the remnant object, which is an active field
of research at the moment.} The no-hair conjecture in GR~\citep{}
states that an (electrically neutral) astrophysical BH is completely
described by two observables: mass and spin angular momentum. One
consequence of the no-hair conjecture is that the (complex) QNM
frequencies of gravitational radiation emitted by a perturbed isolated
BH \ab{are} \sout{is} uniquely determined by its mass and spin angular momentum. Hence
a test of the no-hair conjecture would involve checking for
consistency between estimates of mass and spin of the remnant object
across multiple QNM frequencies~\cite{Dreyer:2003bv}. An inconsistency would either
indicate a non-BH nature of the remnant object, or an incompleteness
of GR as the underlying theory of gravity.

>>>>

\title{Constraints on the quasi-normal mode frequencies of the LIGO-Virgo signals by making full use of gravitational-wave modeling}

@@

The LIGO Scientific Collaboration~\citep{lsc} and the Virgo
Collaboration~\citep{Virgo} have recently announced their catalogue of
gravitational-wave (GW) signals from the first
half of the third observing run (O3a)~\cite{Abbott:2020niy}. Combined
with the first and second observing-run catalogues~\cite{LIGOScientific:2018mvr}, the Advanced LIGO detectors at Hanford,
Washington and Livingston, Louisiana~\citep{aasi2015characterization},
and the Advanced Virgo detector in Cascina,
Italy~\citep{acernese2014advanced} have now detected $50$ GW
events from the merger of compact objects like neutron stars and/or
black holes (BHs). Alongside independent claims of
detections~\citep{nitz20191,nitz20202,2019PhRvD.100b3007Z,2020PhRvD.101h3030V,Venumadhav_2020}, these results have firmly established the field of GW astronomy, five years after the first detection of a GW passing through Earth, notably GW150914~\citep{abbott2016observation}.

The observation of GWs has had significant astrophysical and cosmological
implications~\citep{LSC_2016astroph,gw170817_mma,gw170817_joint,gw170817_hubble}. It
has also allowed us to probe fundamental
physics and test predictions of Einstein's theory of General Relativity 
(GR) in the previously unexplored highly-dynamical, and strong field 
regime~\citep{LSC_2016grtests,GW170817_TGR,gwtc1_tgr}. In GR, a binary black hole (BBH) system is described by
three distinct phases: an early \textit{inspiral}, where the two
compact objects spiral in loosing energy because of the emission of GWs, a \textit{merger} marked by the
formation of a common apparent horizon~\citep{NRpaper}, and a \textit{ringdown}, during which the newly formed remnant object settles down to a Kerr BH emitting quasi-normal-modes (QNMs)~\citep{vishu,earlyqnmpapers} (i.e., damped oscillations with specific, discrete frequencies and decay times).

The LIGO-Virgo collaborations have released companion papers \cite{} detailing their results 
of tests of GR for GW150914 \cite{} and for several GW events of 
the two transient catalogues (TC): GWTC-1\cite{}  and GWTC-2 \cite{}. 
The results include tests of GW generation and source dynamics, where bounds are placed on
parameterized deviations in the post-Newtonian (PN) coefficients describing
the early inspiral~\citep{earlydevelopmentpapers}, and
phenomenological coefficients describing the intermediate (plunge) and
merger regimes of coalescence~\citep{TIGERmethodspapers}; tests of GW
propagation, which assume a generalized dispersion relation and place
upper bounds on the Compton wavelength and, consequently, the mass of
the graviton~\citep{gw170104,samajdar2017projected}, and tests of the
polarization of gravitational radiation using a
multi--GW-detector network~\citep{gw170814,isi2017probing}. The GWTC-1/2 papers also 
check for consistency between different portions of the signal using estimates for the predicted mass and spin of the remnant
object~\citep{Ghosh:2016xx,Ghosh:2017gfp,LSC_2016grtests}, and
consistency of the residuals with interferometric
noise~\citep{Ghonge:2020suv,gwtc1_tgr}. None of these tests report any
departure from the predictions of GR.

The first paper on tests of GR by the LIGO Collaboration~\cite{} provided us with the 
first measurement of the dominant damped-oscillation signal in the ringdown stage of a BBH 
coalescence. The LIGO-Virgo O3a testing GR paper~\cite{} has recently reported a comprehensive analysis of the properties of the remnant object, including the ringdown stage, for tens of GW events.
The no-hair conjecture in GR~\citep{} states that an (electrically neutral) astrophysical BH is completely
described by two observables: mass and spin angular momentum. One
consequence of the no-hair conjecture is that the (complex) QNM
frequencies of gravitational radiation emitted by a perturbed isolated
BH are uniquely determined by its mass and spin angular momentum. Hence
a test of the no-hair conjecture would involve checking for
consistency between estimates of mass and spin of the remnant object
across multiple QNM frequencies~\cite{Dreyer:2003bv}. An inconsistency would either
indicate a non-BH nature of the remnant object, or an incompleteness
of GR as the underlying theory of gravity.
